---
title: "Analysis with Regularization on Brazilian elections"
subtitle: "Predictive analysis with regularization and Hyperparameter tuning on data about Brazilian elections"
author: "José Benardi de Souza Nunes"
date: 02/11/2018
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

# Introduction

<br>

> Data Analysis with multivariate Linear Regression  on data about polls for the 2006 and 2010 elections in Brazil for the "Câmara Federal de Deputados". Data was taken from the [TSE portal](http://www.tse.jus.br/) and encompasses approximately 7300 candidates.

<br>

***

<br>

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(dataPreparation)
library(tidyverse)
library(caret)
library(here)
```

# Data Overview

<br>

## The variables

<br>

```
The response variable is the variable that you are interested in reaching conclusions about.

A predictor variable is a variable used to predict another variable.

Our response variable will be "votos", we want to study how well the predictor variables can help predict its behavior and how they impact in the linear regression.
```

<br>

#### Each item corresponds to a candidate, the attributes of each item are as follows:

- **ano** : Year at which the election took place.
- **sequencial_candidato** : Sequential ID to map the candidates
- **nome** : Name of the candidate
- **uf** : Federate state to which the candidate belongs.
- **partido** : Political party to which the candidate belongs.
- **quantidade_doacoes** : Number of donations received during political campaign.
- **quantidade_doadores** : Number of donators that contributed to the candidate's political campaign.
- **total_receita** : Total revenue. 
- **media_receita** : Mean revenue. 
- **recursos_de_outros_candidatos.comites** : Revenue coming from other candidate's committees.
- **recursos_de_pessoas_fisicas** : Revenue coming from individuals.
- **recursos_de_pessoas_juridicas** : Revenue coming from legal entities.
- **recursos_proprios** : Revenue coming from personal resources.
- **recursos_de_partido_politico** : Revenue coming from political party.
- **quantidade_despesas** : Number of expenses.
- **quantidade_fornecedores** : Number of suppliers.
- **total_despesa** : Total expenditure. 
- **media_despesa** : Mea expenditure.
- **cargo** : Position.
- **sexo** : Sex.
- **grau** : Level of education.
- **estado_civil** : Marital status. 
- **ocupacao** : Candidate's occupation up to the election.
- **votos** : Number of votes received.

<br>

## Loading Data

```{r}
data <- readr::read_csv(
  here::here('data/train.csv'), 
  progress = FALSE,
  local=readr::locale("br"),
  col_types = cols(
    ano = col_integer(),
    sequencial_candidato = col_character(),
    quantidade_doacoes = col_integer(),
    quantidade_doadores = col_integer(),
    total_receita = col_double(),
    media_receita = col_double(),
    recursos_de_outros_candidatos.comites = col_double(),
    recursos_de_pessoas_fisicas = col_double(),
    recursos_de_pessoas_juridicas = col_double(),
    recursos_proprios = col_double(),
    `recursos_de_partido_politico` = col_double(),
    quantidade_despesas = col_integer(),
    quantidade_fornecedores = col_integer(),
    total_despesa = col_double(),
    media_despesa = col_double(),
    votos = col_integer(),
    .default = col_character())) %>%
  mutate(sequencial_candidato = as.numeric(sequencial_candidato),
         estado_civil = as.factor(estado_civil),
         ocupacao = as.factor(ocupacao),
         partido = as.factor(partido),
         grau = as.factor(grau),
         sexo = as.factor(sexo),
         uf = as.factor(uf))

data %>% 
  glimpse()
```

```{r}
data_test <- readr::read_csv(
  here::here('data/test.csv'), 
  progress = FALSE,
  local=readr::locale("br"),
  col_types = cols(
    ano = col_integer(),
    sequencial_candidato = col_character(),
    quantidade_doacoes = col_integer(),
    quantidade_doadores = col_integer(),
    total_receita = col_double(),
    media_receita = col_double(),
    recursos_de_outros_candidatos.comites = col_double(),
    recursos_de_pessoas_fisicas = col_double(),
    recursos_de_pessoas_juridicas = col_double(),
    recursos_proprios = col_double(),
    `recursos_de_partido_politico` = col_double(),
    quantidade_despesas = col_integer(),
    quantidade_fornecedores = col_integer(),
    total_despesa = col_double(),
    media_despesa = col_double(),
    .default = col_character())) %>%
  mutate(sequencial_candidato = as.numeric(sequencial_candidato))

data_test %>% 
  glimpse()
```

<br>

## Assessing data integrity

### NA values

```{r}
data %>%
  map_df(function(x) sum(is.na(x))) %>%
  gather(feature, num_nulls) %>%
  arrange(desc(num_nulls))
```

* No null values found

```{r}
data_test %>%
  map_df(function(x) sum(is.na(x))) %>%
  gather(feature, num_nulls) %>%
  arrange(desc(num_nulls))
```

* No null values found

<br>

### Encoding

We must apply the same encoding to the **data used for the competition test** and the data **we’ll use to build our models** in order to make sure the levels of the categorical variables across both datasets match.

```{r}
encoding <- build_encoding(dataSet = data,
                           cols = c("uf","sexo","grau","ocupacao",
                                    "partido","estado_civil"),
                           verbose = F)

data <- one_hot_encoder(dataSet = data,
                           encoding = encoding,
                           drop = TRUE,
                           verbose = F)

cat("#### Data Shape",
    "\n##### Observations: ",nrow(data),
    "\n##### Variables: ",ncol(data))
```

<br>

```{r}
data_test <- one_hot_encoder(dataSet = data_test,
                           encoding = encoding,
                           drop = TRUE,
                           verbose = F)

cat("#### Test Data Shape",
    "\n##### Observations: ",nrow(data_test),
    "\n##### Variables: ",ncol(data_test))
```

<br>

```{r}
data %>%
  nearZeroVar(saveMetrics = TRUE) %>%
  tibble::rownames_to_column("variable") %>%
  filter(nzv == T) %>% 
  pull(variable) -> near_zero_vars

near_zero_vars %>% 
  glimpse() 
```

* These predictors have a near zero variance, so they behave much like a constant. Predictors that remain constant have no impact on the response variable and for that reason are not useful.

> Following this information we shall exclude predictors of near zero variance and zero variance from our models.

<br>

***

<br>

# Ridge

<br>

* Let’s employ linear regression with regularization through the Ridge method and tune the hyperparameter $\lambda \ (lambda)$

```{r}
set.seed(131)

lambdaGrid <- expand.grid(lambda = 10^seq(10, -2, length=100))

fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv", 
                           number = 10)

data %>%
  select(-one_of(near_zero_vars)) %>%
  select(-ano,-nome) %>%
  train(votos ~ .,
        data = .,
        method = "ridge",
        na.action = na.omit,
        tuneGrid = lambdaGrid,
        trControl = fitControl,
        preProcess = c('scale', 'center')) -> model.ridge

model.ridge
```

* The best $RSME$ value was $35004.01$ and the corresponding best value for the hyperparameter $\lambda$ was $0.01$

```{r}
model.ridge %>%
  varImp() %$%
  importance %>%
  as.data.frame() %>%
  rownames_to_column(var="Feature") %>%
  mutate(Feature = tolower(Feature)) %>%
  ggplot() +
  geom_col(aes(x = reorder(Feature,Overall),
               y = Overall)) + 
  labs(x="Feature", y="Overall Importance") +
  coord_flip()
```

* We have **total_receita, total_despesa** and **recursos_de_pessoas_juridicas** as the three most important features.

* The model ignored features such as **media_despesa, recursos_de_outros_candidatos.comites** and **recursos_proprios**.


<br>

***

<br>

# Lasso

<br>

Let’s employ linear regression with regularization through the Lasso method and tune the hyperparameter $\lambda \ (lambda)$ which in this package is made available as $fraction$.

```{r}
set.seed(131)

fractionGrid <- expand.grid(fraction = seq(1, 1e-2, length=100))

data %>%
  select(-one_of(near_zero_vars)) %>%
  select(-ano,-nome) %>%
  train(votos ~ .,
        data = .,
        method = "lasso",
        na.action = na.omit,
        tuneGrid = fractionGrid,
        trControl = fitControl,
        preProcess = c('scale', 'center')) -> model.lasso

model.lasso
```

* The best $RSME$ value was $35052.40$ and the corresponding best value for the hyperparameter $fraction$ was $0.47$

```{r}
model.lasso %>%
  varImp() %$%
  importance %>%
  as.data.frame() %>%
  rownames_to_column(var="Feature") %>%
  mutate(Feature = tolower(Feature)) %>%
  ggplot() +
  geom_col(aes(x = reorder(Feature,Overall),
               y = Overall)) + 
  labs(x="Feature", y="Overall Importance") +
  coord_flip()
```

* We have **total_receita, total_despesa** and **recursos_de_pessoas_juridicas** as the three most important features.

* The model ignored features such as **media_despesa, recursos_de_outros_candidatos.comites** and **recursos_proprios**.

<br>

***

<br>

# k nearest neighbors

<br>

* Let’s employ the non parametric **k nearest neighbors** regression and tune the hyperparameter 

```{r}
set.seed(131)

neighborsGrid <- expand.grid(k = seq(1, 100, length=100))

data %>%
  select(-one_of(near_zero_vars)) %>%
  select(-ano,-nome) %>%
  train(votos ~ .,
        data = .,
        method = "knn",
        na.action = na.omit,
        tuneGrid = neighborsGrid,
        trControl = fitControl,
        preProcess = c('scale', 'center')) -> model.knn

model.knn
```

* The best $RSME$ value was $34289.06$ and the corresponding best value for the hyperparameter $k$ was $16$

```{r}
model.knn %>%
  varImp() %$%
  importance %>%
  as.data.frame() %>%
  rownames_to_column(var="Feature") %>%
  mutate(Feature = tolower(Feature)) %>%
  ggplot() +
  geom_col(aes(x = reorder(Feature,Overall),
               y = Overall)) + 
  labs(x="Feature", y="Overall Importance") +
  coord_flip()
```

* We have **total_receita, total_despesa** and **recursos_de_pessoas_juridicas** as the three most important features.

* The model paid little attention to features such as **media_despesa, recursos_de_outros_candidatos.comites** and **recursos_proprios**.

<br>

***

<br>

# Comparison between models

<br>

## Importance of Features

<br>

Across the different models there was considerable consensus regarding importance of features. The following statements do encompass both Lasso and Ridge:

* **total_receita, total_despesa** and **recursos_de_pessoas_juridicas** were pointed as the most important features

* **media_despesa, recursos_de_outros_candidatos.comites** and **recursos_proprios** were ignored.

<br>

## Quality Measures (RMSE)

* **Ridge**: RSME = 35004.01
* **Lasso**: RSME = 35052.40
* **KNN**: RSME = 34289.06

>The best performing model was KNN, which will be trained with its optimal hyperparameter set (k = 16).

<br>

***

<br>

# Final model

<br>

```{r}
set.seed(131)

data %>%
  select(-one_of(near_zero_vars)) %>%
  select(-ano,-nome) %>%
  train(votos ~ .,
        data = .,
        method = "knn",
        na.action = na.omit,
        tuneGrid = data.frame(k = 16),
        trControl = trainControl(method="none"),
        preProcess = c('scale', 'center')) -> model.knn.best

model.knn.best
```

```{r}
model.knn.best %>%
  varImp() %$%
  importance %>%
  as.data.frame() %>%
  rownames_to_column(var="Feature") %>%
  mutate(Feature = tolower(Feature)) %>%
  ggplot() +
  geom_col(aes(x = reorder(Feature,Overall),
               y = Overall)) + 
  labs(x="Feature", y="Overall Importance") +
  coord_flip()
```

* We have **total_receita, total_despesa** and **recursos_de_pessoas_juridicas** as the three most important features.

* The model paid little attention to features such as **media_despesa, recursos_de_outros_candidatos.comites** and **recursos_proprios**.

<br>

## Making actual predictions

```{r}
data_test %>%
  mutate(sequencial_candidato = as.character(sequencial_candidato)) %>%
  pull(sequencial_candidato) -> id_column

predict(model.knn.best, data_test) -> predictions

data.frame(ID = id_column,
           votos = predictions) -> submission

submission %>%
  glimpse()
```

```{r}
write_csv(submission,
          here::here('data/submission.csv'))
```


